---
title: "Homework 6"
author: "Rose Killian"
output: github_document
---

### Loading libraries and data

```{r message= FALSE}
library(tidyverse)
library(modelr)
library(mgcv)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r, data import tidy}
birth_df = 
  read_csv("data/birthweight.csv") %>% 
    janitor::clean_names() %>% 
  mutate(
    babysex = recode_factor(babysex, `1` = "male", `2` = "female"),
    frace = recode_factor(frace, `1` = "white", `2` = "black", `3` = "asian", `4` = "puerto rican", `8` = "other", `9` = "unknown"),
    mrace = recode_factor(mrace, `1` = "white", `2` = "black", `3` = "asian", `4` = "puerto rican", `8` = "other", `9` = "unknown"),
    malform = recode_factor(malform, `0` = "absent", `1` = "present")
  )

sum(is.na(birth_df))
```

No missing observations


Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

After doing some general research on known risk-factors of low birth weight I will start to build a model by examining the effects of length at birth, gestational age, sex, and cigarettes smoked on birth weight.

Based on the literature, I would also have included parity, number of previous low birth weight babies, and number of previous small for gestational age babies, but almost every value for parity is zero (total previous births = `r sum(pull(birth_df, parity))` with range `r min(pull(birth_df, parity))` - `r max(pull(birth_df, parity))`) and no observation in the sample has a previous low birth weight or small for gestational age baby.

```{r, model 1}
fit = lm(bwt ~ blength + gaweeks + babysex + smoken, data = birth_df)

fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

birth_df %>% 
  modelr::add_residuals(fit) %>% 
  modelr::add_predictions(fit) %>% 
  ggplot(aes(x = pred, y = resid, alpha = 0.3)) +
  geom_point()
```


Compare your model to two others:

* One using length at birth and gestational age as predictors (main effects only)

```{r, model 2}
fit_2 = lm(bwt ~ blength + gaweeks, data = birth_df)

fit_2 %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

birth_df %>% 
  modelr::add_residuals(fit_2) %>% 
  modelr::add_predictions(fit_2) %>% 
  ggplot(aes(x = pred, y = resid, alpha = 0.3)) +
  geom_point()
  
```


* One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r, model 3}
fit_3 = lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + blength * bhead * babysex, data = birth_df)

fit_3 %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

birth_df %>% 
  modelr::add_residuals(fit_3) %>% 
  modelr::add_predictions(fit_3) %>% 
  ggplot(aes(x = pred, y = resid, alpha = 0.3)) +
  geom_point()
```


Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.

# Problem 2

```{r, data import}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response and tmin as the predictor, and are interested in the distribution of two quantities estimated from these data:

* r^2

* log(β^0∗β^1)

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r^2
and log(β^0∗β^1). Note: broom::glance() is helpful for extracting r^2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β^0∗β^1).